Values disclosures and trust in science: A replication study

### 2022-10-11 paper/paper.Qmd
edit paper + meti subscale correlation
made revisions to Paper.qmd and added code to assess the correlations between the METI subscales in the 02_analysis script

After viewing the slide, the participant is then asked to rate Dr. Spence's trustworthiness using a semantic differential scale.  @ElliottValuesEnvironmentalResearch2017 used an ad hoc scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.  On the METI, participants rate the target scientist (the fictional Dr. Spence) on a 1-7 scale for 14 semantic differential items, where each item is anchored at the ends by a pair of words, such as competent-incompetent or responsible-irresponsible. The METI scale captures perceived trustworthiness of a target along three dimensions: competence, benevolence, and integrity. However, the three dimensions were strongly correlated in our sample. To avoid issues of multicollinearity in our analyses, we averaged participant scores into a single composite measure. In both @ElliottValuesEnvironmentalResearch2017 and our replication analysis, all items were averaged together to form a 1-7 composite measure of trustworthiness.  To aid interpretation, in the analysis the direction of the scale has been set so that increased values correspond to increased perceived trustworthiness.

@@

After filling in the METI, participants provided demographic information and other sections of the survey that are not examined here.  Due to researcher error, a question about the participants' values (whether they prioritize economic growth or public health) was omitted in the first wave of data collection; this question was asked in a followup wave.

@@

H5. Variation in effects
~ The magnitude of the effects 2-4 vary depending on whether the participant prioritizes public health or economic growth.

@@

Hypothesis 1 was analyzed using Spearman rank order correlation and visually.  Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed *a priori* to identify appropriate adjustments (covariates) for H4 and 5.  Because @ElliottValuesEnvironmentalResearch2017 made their data publicly available, and our analytical approaches differ somewhat from theirs, we conducted parallel analyses of their data, and report these parallel results for H2 and 3.  Exploratory data analysis was used throughout to support data validation and aid interpretation.

@@

After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample ($M_{age}$ = 44-years-old, $SD_{age}$ = 16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124, Asian or Pacific Islander = 63, Hispanic = 33, American Indian or Alaskan Native = 5, Mixed or Other = 51).  Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition (@tbl-condition). Of the full 988 participants, 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).  Consequently, subsamples for hypotheses 4 and 5 were substantially smaller than the full analysis sample.

@@

The study was approved by the UC Merced IRB on August 17, 2021, and data collection ran October 18-20, 2021.  As mentioned above, due to researcher error a question about participants' values was not included in the original survey.  A followup survey asking this question of the same group of participants was conducted December 8, 2021 through March 5, 2022.

>>>>

After viewing the slide, the participant is then asked to rate Dr. Spence's trustworthiness using a semantic differential scale.  @ElliottValuesEnvironmentalResearch2017 used an ad hoc scale; we used the Muenster Epistemic Trustworthiness Inventory [METI\; @HendriksMeasuringLaypeopleTrust2015], which substantially overlaps but is not identical to the @ElliottValuesEnvironmentalResearch2017 scale.  Even though the ad hoc scale created by Elliot and colleagues had acceptable internal reliability, the METI was developed and psychometrically validated to measure the perceived trustworthiness of experts. On the METI, participants rate the target scientist (the fictional Dr. Spence) on a 1-7 scale for 14 semantic differential items, where each item is anchored at the ends by a pair of words, such as competent-incompetent or responsible-irresponsible. The METI scale captures perceived trustworthiness of a target along three dimensions: competence, benevolence, and integrity. However, the three dimensions were strongly correlated in our sample (Pearson r values ranging from .75 to .92). To avoid issues of multicollinearity in our analyses, we averaged participant scores into a single composite measure. In both @ElliottValuesEnvironmentalResearch2017 and our replication analysis, all items were averaged together to form a 1-7 composite measure of trustworthiness.  To aid interpretation, in the analysis the direction of the scale has been set so that increased values correspond to increased perceived trustworthiness.

@@

After filling in the METI, participants provided demographic information, including self-identifying their political ideology, and other sections of the survey that are not examined here.  Due to researcher error, a question about the participants' values (whether they prioritize economic growth or public health) was omitted in the first wave of data collection; this question was asked in a followup wave.

@@

H5. Variation in effects
~ The magnitude of the effects for hypotheses 2-4 vary depending on whether the participant prioritizes public health or economic growth.

@@

Hypothesis 1 was analyzed using Spearman rank order correlation to test H1a and a visual inspection of descriptive statistics to test H1b.  Hypotheses 2-5 were analyzed using linear regression models as a common framework, with a direct acyclic graph (DAG) constructed *a priori* to identify appropriate adjustments (covariates) for H4 and 5.  Because @ElliottValuesEnvironmentalResearch2017 made their data publicly available, and our analytical approaches differ somewhat from theirs, we conducted parallel analyses of their data, and report these parallel results for H2 and 3.  Exploratory data analysis was used throughout to support data validation and aid interpretation.

@@

After excluding participants who declined consent after opening the survey or did not complete the survey, we had 988 participants in the full analysis sample ($M_{age}$ = 44-years-old, $SD_{age}$ = 16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124, Asian or Pacific Islander = 63, Hispanic = 33, American Indian or Alaskan Native = 5, Mixed or Other = 51).  Participants were randomly assigned to condition, with 163 assigned to the No Disclosure + Causes Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm condition, 165 assigned to the Economic Growth + Causes Harm condition, 165 assigned to the Economic Growth + Does Not Cause Harm condition, 168 assigned to the Public Health + Causes Harm condition, and 162 assigned to the Public Health + Does Not Cause Harm condition (@tbl-condition). Due to researcher error a question about participants' values was not included in the original survey. Of the full 988 participants, 844 participants (85%) responded to the followup question about their own values (participant prioritizes economic growth or public health).  Consequently, subsamples for hypotheses 4 and 5 were substantially smaller than the full analysis sample.

@@

The study was approved by the UC Merced IRB on August 17, 2021, and data collection ran October 18-20, 2021. The followup survey asking the initial sample of participants about their own values regarding economic growth and public health was conducted December 8, 2021 through March 5, 2022.
